### 2025-04-04 GettingStarted.Rmd
install and test with vignettes

---
title: "WHAM MSE startup"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Following the vignettes from Chengxue Li at https://lichengxue.github.io/whamMSE/

Installed as instructed under https://lichengxue.github.io/whamMSE/01.Installation.html but WHAM dev branch already installed

Running self test and cross test https://lichengxue.github.io/whamMSE/02.Self-test-and-Cross-test.html

Simple self test need to change some inputs to avoid error, code modified from vignette below, all other code blocks from vignette

```{r}
year_start  <- 1  # starting year in the burn-in period
year_end    <- 20  # end year in the burn-in period
MSE_years   <- 0     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation

info <- generate_basic_info(n_stocks = 2,
                            n_regions = 2,
                            n_indices = 2,
                            n_fleets = 2,
                            n_seasons = 4,
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "medium",
                            n_ages = 12,
                            #Fbar_ages = 12,
                            #recruit_model = 2,
                            F_info = list(F.year1 = 0.2, Fhist = "constant", Fmax = 0.4, Fmin = 0.2, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = 0.625, q = 0.2),
                            fracyr_spawn = 0.625,
                            fracyr_seasons = NULL,
                            fleet_regions = NULL, # was fleet_pointer
                            index_regions = NULL, # was index_pointer
                            user_waa = NULL,
                            user_maturity = NULL,
                            bias.correct.process = FALSE,
                            bias.correct.observation = FALSE,
                            bias.correct.BRPs = FALSE,
                            mig_type = 0,
                            XSPR_R_opt = 2,
                            move_dyn = 0, # natal homing is assumed
                            onto_move = 0,
                            onto_move_pars = NULL,
                            apply_re_trend = 0,
                            trend_re_rate = NULL,
                            apply_mu_trend = 0,
                            trend_mu_rate = NULL, 
                            age_mu_devs = NULL)

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info
```

Which option to run? Trying 3a bi-directional

```{r}
basic_info <- generate_NAA_where(basic_info = basic_info, move.type = 2)
# Note: the default is "bidirectional" movement (e.g. stock 1 move to region 2 and stock 2 move to region 1)

move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3, 0.1), move.re = "constant") # Note: default is move = 0.3 (constant) for stock1 and 0.1 (constant) for the other stocks
```

Selectivity and M, 0.2 of course

```{r}
n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
```

NAA

```{r}
sigma        <- "rec+1"
re_cor       <- "iid"
ini.opt      <- "equilibrium" # option <- c("age-specific-fe", "equilibrium")
Rec_sig      <- 0.5 # (sigma for recruitment)
NAA_sig      <- 0.5 # (sigma for NAA)

sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages
sigma_vals[,,1] = Rec_sig

# Set initial NAA for each stock
log_N1    <- rep(10,n_stocks)
log_N1[1] <- log(exp(10)*2) # N1_stock1 is 2 times higher than N1_stock2
N1_pars   <- generate_ini_N1(log_N1,basic_info,ini.opt)

# Set mean recruitment para. for each stock
mean_rec_par <- list()
for (i in 1:n_stocks) mean_rec_par[[i]] <- exp(log_N1[i])

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 2,  # rec random around the mean
               recruit_pars = mean_rec_par, 
               sigma_vals = sigma_vals,  
               N1_pars = N1_pars,
               NAA_where = basic_info$NAA_where) # NAA_where must be specified in basic_info!
```

WHAM input

```{r}
input <- prepare_wham_input(basic_info = basic_info, 
                            selectivity = sel, 
                            M = M, 
                            NAA_re = NAA_re, 
                            move = move,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)
```

OM generation

```{r}
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)
```

Self test

```{r}
sim_fn <- function(om, self.fit = FALSE){
  input <- om$input
  input$data <- om$simulate(complete=TRUE)
  # $simulate() is a function to generate datasets given the parameters defined in the OM
  # Note: process error and observation error are both turned on when simulating datasets 
  # To turn them off, use basic_info$simulate_process_error = 0 and basic_info$simulate_observation_error = 0
  if(self.fit) {
    fit <- fit_wham(input, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = FALSE)
    return(fit)
  } else return(input) 
}
# Note: turn on do.osa to calculate one-step-ahead residuals
# Note: turn on do.retro to calculate retrospective bias
set.seed(12345)
self_sim_fit <- sim_fn(om, self.fit = TRUE)
```

Check convergence (pass)

```{r}
check_convergence(self_sim_fit) # check the model convergence
```

Cross test: setup

```{r}
# EM with different NAA configuration
sigma      <- "rec+1"
re_cor     <- "2dar1" # Misspecification here

ini.opt    <- "equilibrium"
NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               NAA_where = basic_info$NAA_where)

# Generate wham input
input <- prepare_wham_input(basic_info = basic_info, selectivity = sel, M = M, NAA_re = NAA_re, move = move,
                            catch_info = catch_info, index_info = index_info, F = F_info)

# Generate EM
em <- fit_wham(input, do.fit = F, do.brps = F, MakeADFun.silent = TRUE)
```

Cross test: perform

```{r}
sim_fn2 <- function(om, em, cross.fit = FALSE){
  input <- em$input
  input$data <- om$simulate(complete=TRUE)
  if(cross.fit) {
    fit <- fit_wham(input, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = FALSE)
    return(fit)
  } else return(input) 
}

set.seed(12345)
cross_sim_fit <- sim_fn2(om, em, cross.fit = TRUE)
```

Cross test: convergence test (fails)

```{r}
check_convergence(cross_sim_fit)
```
Replicates

Self test

```{r}
nsim <- 100 
set.seed(8675309) 
sim_input <- list()
sim_input <- lapply(1:nsim, function(x) {
  input_i <- om$input
  sim <- om$simulate(complete=TRUE)
  input_i$data <- sim
  return(input_i)
})

# Self test
sim_fits <- list()
sim_fits <- lapply(1:nsim, function(x){
  cat(paste("model_fit:", x, "start \n"))
  out <- fit_wham(sim_input[[x]], do.osa = FALSE, MakeADFun.silent = TRUE, retro.silent = TRUE, save.sdrep = FALSE)
  cat(paste("model_fit:", x, "done \n"))
  return(out)
})

# Summarize results 
conv <- sapply(1:nsim, function(x){
  if (length(sim_fits[[x]]) != 0) {
    if (sim_fits[[x]]$is_sdrep & !sim_fits[[x]]$na_sdrep & !sim_fits[[x]]$hessian) {
      conv = TRUE } else conv = FALSE
  } else conv = FALSE
  return(conv)
})
cat(paste("Convergence rate:", sum(conv)/nsim))

mean_rec_par <- lapply(1:nsim, function(x){
  mean_rec_par_est <- sim_fits[[x]]$parList$mean_rec_pars[,1]
  mean_rec_par_true <- sim_fits[[x]]$input$par$mean_rec_pars[,1]
  mean_rec_par <- cbind(mean_rec_par_est,mean_rec_par_true)
  return(mean_rec_par)
})
print(mean_rec_par)

SSB <- lapply(1:nsim, function(x){
  SSB_est <- sim_fits[[x]]$rep$SSB
  SSB_true <- sim_fits[[x]]$input$data$SSB
  SSB <- cbind(SSB_est, SSB_true)
  return(SSB)
})
print(SSB)
```