### 2025-04-07 docs/index.Rmd
start pages

---
title: "WHAM MSE Pages"
author: "Sarah Gaichas"
output: html_document
---

<button>
    <a href="https://github.com/NOAA-EDAB/WHAMmse_longlived"> WHAM mse longlived github repo </a>
</button>

<!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

## Analysis pages

<a href="https://noaa-edab.github.io/WHAMmse_longlived/GettingStarted.html" target="_blank"> Test WHAM MSE code from vignettes - April 2025 </a>

<a href="https://noaa-edab.github.io/WHAMmse_longlived/WorkflowDecisions.html" target="_blank"> Documenting long lived MSE workflow - April 2025 </a>

<a href="https://noaa-edab.github.io/WHAMmse_longlived/Results.html" target="_blank"> Initial results - April 2025 </a>