### 2025-04-07 MSEsetup.Rmd
trying long lived MSE setup, currently breaking

---
title: "MSE Setup"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output:
  html_document:
    code_fold: hide
link-citations: yes
csl: "canadian-journal-of-fisheries-and-aquatic-sciences.csl"
bibliography: MSE.bib
urlcolor: blue  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

Following the MSE vignette from Chengxue Li at https://lichengxue.github.io/whamMSE/

```{r}
library(wham)
library(whamMSE)

main.dir = here::here()
```

## Run the vignette exactly first, breaks

```{r, eval=FALSE}
year_start  <- 1  # starting year in the burn-in period
year_end    <- 20  # end year in the burn-in period
MSE_years   <- 3     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation

info <- generate_basic_info(n_stocks = 2,
                            n_regions = 2,
                            n_indices = 2,
                            n_fleets = 2,
                            n_seasons = 4,
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "medium",
                            n_ages = 12,
                            #Fbar_ages = 12,
                            #recruit_model = 2,
                            F_info = list(F.year1 = 0.2, Fhist = "F-H-L", Fmax = 2, Fmin = 1, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = 0.625, q = 0.2),
                            fracyr_spawn = 0.625)

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info

basic_info <- generate_NAA_where(basic_info = basic_info, move.type = 2) # "bidirectional" movement

move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3,0.1), 
                      move.re = "iid_y", move.sigma = 0.5, prior.sigma = 0.5, use.prior = TRUE)

n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))

sigma      <- "rec+1"
re_cor     <- "iid"
ini.opt    <- "equilibrium" # option   <- c("age-specific-fe", "equilibrium")

# Set para. for B-H function
alpha <- 12
beta  <- 1.5e-4

# Set sigma for NAA
NAA_sig <- 0.2
sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"

# Set initial NAA for each stock
log_N1  <- rep(10, n_stocks) # Create difference between stocks
N1_pars <- generate_ini_N1(log_N1,basic_info,ini.opt)

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 3,
               recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
               sigma_vals = sigma_vals,
               N1_pars = N1_pars,
               NAA_where = basic_info$NAA_where)

# recruit_model = 1: estimating annual recruitments as fixed effects or a random walk if NAA_re$sigma specified
# recruit_model = 2: estimating a mean recruitment with annual recruitments as random effects
# recruit_model = 3: Beverton-Holt stock-recruitment with annual recruitments as random effects
# recruit_model = 4: Ricker stock-recruitment with annual recruitments as random effects

# 1. recruit_pars: a list (length = n_stocks) of vectors of initial parameters for recruitment model. 
# If $recruit_model is 3 (B-H) or 4 (Ricker), parameters are "alpha" and "beta".

# 2. sigma_vals: Initial standard deviation values to use for the NAA deviations. Values are not used if recruit_model = 1 
# If sigma="rec": must be a list (length = n_stocks) of single values
# If sigma="rec+1": a list (length = n_stocks) of 2 values must be specified. First is for the first age class (recruits), second is for all other ages.

input <- prepare_wham_input(basic_info = basic_info, 
                            selectivity = sel, 
                            M = M, 
                            NAA_re = NAA_re, 
                            move = move,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)

random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)

om_with_data <- update_om_fn(om, seed = 123, random = random)

assess.interval <- 3 # Note: assessment interval is 3 years, given the feedback period is 3 years, there will be only 1 assessment
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs

n_stocks = n_regions = n_fleets = n_indices = 1

sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = TRUE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            aggregate_index_info = list(index_pointer = rep(1,2),
                                                        use_catch_weighted_waa = TRUE),
                            aggregate_catch_info = list(fleet_pointer = rep(1,2),
                                                        use_catch_weighted_waa = TRUE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            hcr = list(hcr.type = 1, hcr.opts = NULL), 
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE)
```

## Modify for long-lived, this runs but crashes

Basic info for a long-lived species. Based on @wiedenmann_evaluation_2017, Table 3 B and C.

A long-lived species recruits at 5 years, lives to 20, mean M is 0.1, matures at 7.

Modify from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html

Start with a 20 year burn in and 20 MSE years. First try constant low F. Then work up to complex case.

Fishing pattern is "updown" increasing to Fmax then decreasing back to Fmin

```{r}

year_start  <- 1  # starting year in the burn-in period
year_end    <- 20  # end year in the burn-in period
MSE_years   <- 20     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation

info <- generate_basic_info(n_stocks = 1, #2,
                            n_regions = 1, #2,
                            n_indices = 1, #2,
                            n_fleets = 1, #2,
                            n_seasons = 4,
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "long",
                            n_ages = 20,
                            #Fbar_ages = 12,
                            #recruit_model = 2,
                            F_info = list(F.year1 = 0.01, Fhist = "constant", Fmax = 2, Fmin = 0.5, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = 0.625, q = 0.2),
                            fracyr_spawn = 0.625)

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info
```

Movement, do we need this? 1 stock 1 region model, turn it off

```{r}
basic_info <- generate_NAA_where(basic_info = basic_info, move.type = NULL) # "bidirectional" movement

move <- NULL

#move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3), #,0.1
#                     move.re = "iid_y", move.sigma = 0.5, prior.sigma = 0.5, use.prior = TRUE)
```
Adjust for 1 area 1 stock, long lived M=0.1

```{r}
n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages)))
```

```{r}
sigma      <- "rec+1"
re_cor     <- "iid"
ini.opt    <- "equilibrium" # option   <- c("age-specific-fe", "equilibrium")

# Set para. for B-H function
alpha <- 12
beta  <- 1.5e-4

# Set sigma for NAA
NAA_sig <- 0.2
sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"

# Set initial NAA for each stock
log_N1  <- rep(10, n_stocks) # Create difference between stocks
N1_pars <- generate_ini_N1(log_N1,basic_info,ini.opt)

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 3,
               recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
               sigma_vals = sigma_vals,
               N1_pars = N1_pars,
               NAA_where = basic_info$NAA_where)

# recruit_model = 1: estimating annual recruitments as fixed effects or a random walk if NAA_re$sigma specified
# recruit_model = 2: estimating a mean recruitment with annual recruitments as random effects
# recruit_model = 3: Beverton-Holt stock-recruitment with annual recruitments as random effects
# recruit_model = 4: Ricker stock-recruitment with annual recruitments as random effects

# 1. recruit_pars: a list (length = n_stocks) of vectors of initial parameters for recruitment model. 
# If $recruit_model is 3 (B-H) or 4 (Ricker), parameters are "alpha" and "beta".

# 2. sigma_vals: Initial standard deviation values to use for the NAA deviations. Values are not used if recruit_model = 1 
# If sigma="rec": must be a list (length = n_stocks) of single values
# If sigma="rec+1": a list (length = n_stocks) of 2 values must be specified. First is for the first age class (recruits), second is for all other ages.
```

Make inputs for the OM

```{r}
 input <- prepare_wham_input(basic_info = basic_info, 
                            selectivity = sel, 
                            M = M, 
                            NAA_re = NAA_re, 
                            move = move,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)
```

Make the OM

```{r}
random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)
```

One dataset

```{r}
om_with_data <- update_om_fn(om, seed = 123, random = random)
```

Assessment specifications, every 3 years as a start, MP could be annual

```{r}
assess.interval <- 3 # 
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs
```

Try a simple EM, looks like EM1 [here](https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#10_Description_of_assessment_models:)

```{r}
# this matches the OM
n_stocks = n_regions = n_fleets = n_indices = 1

sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

M_em <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages))) # this is misspecified M

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = TRUE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            aggregate_index_info = list(index_pointer = rep(1,1),
                                                        use_catch_weighted_waa = TRUE),
                            aggregate_catch_info = list(fleet_pointer = rep(1,1),
                                                        use_catch_weighted_waa = TRUE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            hcr = list(hcr.type = 1, hcr.opts = NULL), 
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE)
```

# References