### 2025-04-11 MSEsetup.Rmd
with Cheng's fixes and additional no survey run

---
title: "MSE Setup"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output:
  html_document:
    code_fold: hide
link-citations: yes
csl: "canadian-journal-of-fisheries-and-aquatic-sciences.csl"
bibliography: MSE.bib
urlcolor: blue  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

Following the MSE vignette from Chengxue Li at https://lichengxue.github.io/whamMSE/

```{r}
library(wham)
library(whamMSE)

main.dir = here::here()
```

## Run the vignette exactly first, Cheng fixed! works now

```{r, eval=FALSE}
year_start  <- 1  # starting year in the burn-in period
year_end    <- 20  # end year in the burn-in period
MSE_years   <- 3     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation

info <- generate_basic_info(n_stocks = 2,
                            n_regions = 2,
                            n_indices = 2,
                            n_fleets = 2,
                            n_seasons = 4,
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "medium",
                            n_ages = 12,
                            #Fbar_ages = 12,
                            #recruit_model = 2,
                            F_info = list(F.year1 = 0.2, Fhist = "F-H-L", Fmax = 2, Fmin = 1, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = 0.625, q = 0.2),
                            fracyr_spawn = 0.625)

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info

basic_info <- generate_NAA_where(basic_info = basic_info, move.type = 2) # "bidirectional" movement

move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3,0.1), 
                      move.re = "iid_y", move.sigma = 0.5, prior.sigma = 0.5, use.prior = TRUE)

n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))

sigma      <- "rec+1"
re_cor     <- "iid"
ini.opt    <- "equilibrium" # option   <- c("age-specific-fe", "equilibrium")

# Set para. for B-H function
alpha <- 12
beta  <- 1.5e-4

# Set sigma for NAA
NAA_sig <- 0.2
sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"

# Set initial NAA for each stock
log_N1  <- rep(10, n_stocks) # Create difference between stocks
N1_pars <- generate_ini_N1(log_N1,basic_info,ini.opt)

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 3,
               recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
               sigma_vals = sigma_vals,
               N1_pars = N1_pars)#,
               #NAA_where = basic_info$NAA_where)

# recruit_model = 1: estimating annual recruitments as fixed effects or a random walk if NAA_re$sigma specified
# recruit_model = 2: estimating a mean recruitment with annual recruitments as random effects
# recruit_model = 3: Beverton-Holt stock-recruitment with annual recruitments as random effects
# recruit_model = 4: Ricker stock-recruitment with annual recruitments as random effects

# 1. recruit_pars: a list (length = n_stocks) of vectors of initial parameters for recruitment model. 
# If $recruit_model is 3 (B-H) or 4 (Ricker), parameters are "alpha" and "beta".

# 2. sigma_vals: Initial standard deviation values to use for the NAA deviations. Values are not used if recruit_model = 1 
# If sigma="rec": must be a list (length = n_stocks) of single values
# If sigma="rec+1": a list (length = n_stocks) of 2 values must be specified. First is for the first age class (recruits), second is for all other ages.

input <- prepare_wham_input(basic_info = basic_info, 
                            selectivity = sel, 
                            M = M, 
                            NAA_re = NAA_re, 
                            move = move,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)

random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)

om_with_data <- update_om_fn(om, seed = 123, random = random)

assess.interval <- 3 # Note: assessment interval is 3 years, given the feedback period is 3 years, there will be only 1 assessment
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs

n_stocks = n_regions = n_fleets = n_indices = 1

sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = TRUE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            # Please use this code:
                            aggregate_catch_info = list(n_fleets = 1, 
                                                        fleet_pointer = c(1,1), 
                                                        use_catch_weighted_waa=TRUE, 
                                                        catch_Neff = 100, 
                                                        catch_cv = 0.1),
                            # Please use this code:
                            aggregate_index_info = list(n_indices = 1, 
                                                        index_pointer = c(1,1), 
                                                        use_catch_weighted_waa=TRUE, 
                                                        index_Neff = 100, 
                                                        index_cv = 0.1),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            hcr = list(hcr.type = 1, hcr.opts = NULL), 
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE)

mods[[1]]$om$rep$pred_catch # Check catch in the feedback loop
mods[[1]]$om$rep$SSB  # Check SSB in the feedback loop
```

## Modify for long-lived, Cheng fixed and now runs

Basic info for a long-lived species. Based on @wiedenmann_evaluation_2017, Table 3 B and C.

A long-lived species recruits at 5 years, lives to 20, mean M is 0.1, matures at 7.

Modify from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html

Start with a 20 year burn in and 20 MSE years. First try constant low F. Then work up to complex case.

Fishing pattern is "constant" at first to get running

Now set up as in new Single Stock MSE vignette: https://lichengxue.github.io/whamMSE/11.Single-Stock-MSE.html

```{r}

year_start  <- 1  # starting year in the burn-in period
year_end    <- 20  # end year in the burn-in period
MSE_years   <- 20     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation

info <- generate_basic_info(n_stocks = 1, #2,
                            n_regions = 1, #2,
                            n_indices = 1, #2,
                            n_fleets = 1, #2,
                            n_seasons = 1, # don't need seasons
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "long",
                            n_ages = 20,
                            # We want a bit higher F in the historical period
                            F_info = list(F.year1 = 0.1, Fhist = "constant", Fmax = 0.1, Fmin = 0.1, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = 0.625, q = 0.2),
                            fracyr_spawn = 0.625)

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info
```

Movement, do we need this? 1 stock 1 region model, turn it off

```{r}
basic_info <- generate_NAA_where(basic_info = basic_info, move.type = NULL) # "bidirectional" movement

move <- NULL

#move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3), #,0.1
#                     move.re = "iid_y", move.sigma = 0.5, prior.sigma = 0.5, use.prior = TRUE)
```
Adjust for 1 area 1 stock, long lived M=0.1

```{r}
n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages)))
```

```{r}
sigma      <- "rec+1"
re_cor     <- "iid"
ini.opt    <- "equilibrium" # option   <- c("age-specific-fe", "equilibrium")

# Set para. for B-H function
alpha <- 12
beta  <- 1.5e-4

# Set sigma for NAA
NAA_sig <- 0.2
sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"

# Set initial NAA for each stock
log_N1  <- rep(10, n_stocks) # Create difference between stocks
N1_pars <- generate_ini_N1(log_N1,basic_info,ini.opt)

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 3,
               recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
               sigma_vals = sigma_vals,
               N1_pars = N1_pars)#,
               #NAA_where = basic_info$NAA_where)

# recruit_model = 1: estimating annual recruitments as fixed effects or a random walk if NAA_re$sigma specified
# recruit_model = 2: estimating a mean recruitment with annual recruitments as random effects
# recruit_model = 3: Beverton-Holt stock-recruitment with annual recruitments as random effects
# recruit_model = 4: Ricker stock-recruitment with annual recruitments as random effects

# 1. recruit_pars: a list (length = n_stocks) of vectors of initial parameters for recruitment model. 
# If $recruit_model is 3 (B-H) or 4 (Ricker), parameters are "alpha" and "beta".

# 2. sigma_vals: Initial standard deviation values to use for the NAA deviations. Values are not used if recruit_model = 1 
# If sigma="rec": must be a list (length = n_stocks) of single values
# If sigma="rec+1": a list (length = n_stocks) of 2 values must be specified. First is for the first age class (recruits), second is for all other ages.
```

Make inputs for the OM

```{r}
 input <- prepare_wham_input(basic_info = basic_info, 
                            selectivity = sel, 
                            M = M, 
                            NAA_re = NAA_re, 
                            move = move,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)

# IMPORTANT!#
# This appears to be due to the initial F value for the Newton iterations to find F40. The default value is too high for long-lived fish
input$data$FXSPR_init[] <- 0.01 # change to a low value

```

Make the OM

```{r}
random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)
```

One dataset

```{r}
om_with_data <- update_om_fn(om, seed = 123, random = random)
```

Assessment specifications, every 3 years as a start, MP could be annual

```{r}
assess.interval <- 3 # 
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs
```

Try a simple EM, looks like EM1 [here](https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#10_Description_of_assessment_models:)

```{r}
# this matches the OM
# n_stocks = n_regions = n_fleets = n_indices = 1
# 
# sel_em <- list(model=rep("logistic",n_fleets+n_indices),
#                initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
# 
NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
# 
# M_em <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages))) # this is misspecified M

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!
```

Try same model with poorer survey in the future

See https://lichengxue.github.io/whamMSE/15.Reduce-Data-Availability-In-Assessment.html#6_Increase_observation_error_and_reduce_sample_size_for_survey_indices_in_the_feedback_period

```{r}
agg_index_sigma = input$data$agg_index_sigma
agg_index_sigma[21:40,] = 1 # Increase CV for both survey indices in the feedback period
index_Neff = input$data$index_Neff
index_Neff[21:40,] = 30 # Decrease ESS for both survey indices in the feedback period

input <- update_input_index_info(input, agg_index_sigma, index_Neff) # Update input file
```

Generate new om and dataset, keep same assessment interval as above

```{r}
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
om_with_data <- update_om_fn(om, seed = 123, random = random)
```

Run new loop

```{r}
mods[[2]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            add.years = TRUE,
                            update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

```

Visualize 2 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)

This compares a good survey in the future (1) with a poor survey in the future (2)

```{r}

par(mfrow = c(1,2))
SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(0,100000), main = "OM SSB") 
colors <- c("red","green","purple","orange")
for (i in length(SSB_s1)) {
  lines(SSB_s1[[i]], col = colors[i-1],lty = i)
}
legend("topleft",legend = paste0("EM ", 1:2), col = c("blue",colors),lty=1:5, cex=0.8)

Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,30000), main = "OM Catch")
colors <- c("red","green","purple","orange")
for (i in length(Catch_s1)) {
  lines(Catch_s1[[i]], col = colors[i-1],lty = i)
}

```
Now add the one with no survey, from Cheng's code

```{r}
# ---------------------------------
# Actually the above section (CV and Neff) will be useless because we will not have survey data in the feedback period), see code below:
# ---------------------------------
remove_agg = TRUE # remove a aggregate index for some years
remove_agg_pointer = 1 # target on index 1
remove_agg_years = 21:40 # all feedback years
remove_paa = TRUE # Also remove age comp for that index 
remove_paa_pointer = 1 # target on index 1
remove_paa_years = 21:40 # all feedback years

input <- update_input_index_info(input, agg_index_sigma, index_Neff,
                                 remove_agg, remove_agg_pointer, remove_agg_years,
                                 remove_paa, remove_paa_pointer, remove_paa_years) # Update input file
```

Generate new om and dataset, keep same assessment interval as above

```{r}
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
om_with_data <- update_om_fn(om, seed = 123, random = random)
```

Run new loop

```{r}
# this simulation actually represents the data-missing scenario.
mods[[3]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            add.years = TRUE,
                            # Must have the below code! Otherwise the EM will assume a constant CV and Neff (same as the CV and Neff in the first year) for the survey
                            update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

```

Visualize 3 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)

```{r}

par(mfrow = c(1,2))
SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(1e4,130000), main = "OM SSB",, lwd = 2) 
lines(SSB_s1[[2]], col = "red",lty = 1, lwd = 2)
lines(SSB_s1[[3]], col = "orange",lty = 1, lwd = 2)

legend("topleft",legend = c("EM1_good_survey", "EM2_poor_survey","EM3_missing_survey"), col = c("blue","red","orange"),lty=1, cex=0.8)

Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,15000), main = "OM Catch", lwd = 2)
lines(Catch_s1[[2]], col = "red",lty = 1, lwd = 2)
lines(Catch_s1[[3]], col = "orange",lty = 1, lwd = 2)

```

# References