### 2025-04-14 Scenarios.Rmd
scenarios

---
title: "MSE Scenarios for short and long lived"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup for MSEs

Single area

Two surveys spring and fall

Testing alternating spring and fall

Baseline is all data all the time,

CV = empirical value from stockeff across short and long lived species

Eff = 100

Catch CV = 10%

Eff = 100

Project at F40%

Assessment frequency 3 years

IID NAA

30 year spin up 30 year loop (10 assessment cycles)

F scenario for spin up? high followed by lower? Yes

"F-H-L" Fmax = 2.5, Fmin = 1, changepoint = 0.5

sigma squared for NAA and recruitment?

cod assessment or redfish for these

## Degraded data quality scenarios:

Alternate spring and fall, double CV and half Eff (50)

Fishery data, double CV (0.2) and half Eff (50)

Keep multinomial

Single fleet logistic selectivity: shift one age to the right from maturity ogive

Different selectivity by spring and fall survey: mimic some long lived species

## Empirical management procedure

Index based method:  Ismooth averaging most recent 3 years of whatever surveys (rescaled to mean of each using Ismooth package)

Ismooth package is here https://github.com/cmlegault/PlanBsmooth

100 replicates for each, try the parallel processing

* Base scenario 
* Alternating surveys and less age comps with 3 year assessments
* Empirical management procedures (later)