### 2025-04-28 PlotScenarios.Rmd
test plots

---
title: "Plot Scenarios"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Test plots with small subset

The first 5 runs of baseline with 3 year assessment are in Results_test

What should we plot? SSB, Catch, F, Rec? How often the stock is overfished, interannual variation in catch?

Read them in and build an object to plot

```{r}
# use the purrr library

# function to pull the pieces from a single rds file

d.name <- here::here("Results_test")
modnames <- list.files(d.name)

modpaths <- paste0(d.name, "/", modnames)

getmodts <- function(path){

  mod <- readRDS(here::here(path))
  modpath <- stringr::str_split(path, "/", simplify = TRUE)
  filename <- modpath[length(modpath)]
  modname <- stringr::str_split(filename, ".RDS", simplify = TRUE)[1]
  sceniter <- stringr::str_split(modname, "_", simplify = TRUE)
  scen <- sceniter[1]
  iter <- sceniter[2]

  omSSB <- mod$om$rep$SSB
  omcatch <- mod$om$rep$pred_catch

   # return model indices as a dataframe
  out <- data.frame(modname = modname,
                    scen = scen,
                    iter = iter,
                    year = 1:length(omSSB),
                    omSSB = omSSB,
                    omcatch = omcatch
                    )

  return(out)
}

# apply the function over all rds in the directory

testdat <- purrr::map_dfr(modpaths, getmodts)

# plot SSB and catch for each model iteration
# needs patchwork library

library(patchwork)

SSB <- ggplot2::ggplot(testdat) +
  ggplot2::geom_line(ggplot2::aes(x=year, y=omSSB, colour = iter)) +
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~scen) +
  ggplot2::theme(legend.position = "none")

Catch <- ggplot2::ggplot(testdat) +
  ggplot2::geom_line(ggplot2::aes(x=year, y=omcatch, colour = iter))+
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~scen) +
  ggplot2::theme(legend.position = "none")

SSB + Catch

```

## A test with all 400

Its a start

```{r}
d.name <- here::here("Results")
modnames <- list.files(d.name)

modpaths <- paste0(d.name, "/", modnames)

# apply the function over all rds in the directory

realdat <- purrr::map_dfr(modpaths, getmodts)

# plot SSB and catch for each model iteration
# needs patchwork library

library(patchwork)

scennames <- c(`Mod1` = "Base 3 yr",
               `Mod2` = "Base 6 yr",
               `Mod3` = "Degraded data 3 yr",
               `Mod4` = "Degraded data 6 yr")

SSB <- ggplot2::ggplot(realdat) +
  ggplot2::geom_line(ggplot2::aes(x=year, y=omSSB, colour = iter),  alpha = 0.3) + #
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~scen, labeller = ggplot2::as_labeller(scennames)) +
  ggplot2::theme(legend.position = "none")

Catch <- ggplot2::ggplot(realdat) +
  ggplot2::geom_line(ggplot2::aes(x=year, y=omcatch, colour = iter), alpha = 0.3) + #
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~scen, labeller = ggplot2::as_labeller(scennames)) +
  ggplot2::theme(legend.position = "none")

SSB

Catch
```